A performance assessment of evolutionary algorithms in volunteer computing environments: the importance of entropy

### 2016-11-14 EvoPar2017/paper.Rnw
Starting with conclusions

Our main intention in this paper was to configure an evolutionary
algorithm in a volunteer computing environment to improve its
efficiency. This initial configuration included
{\em overlaps} or clients running islands that had actually started in
a previous experiment in the set. If we are interested in measuring
how many users or their behavior in bulk this is not so important, but
in this case our focus was on the algorithm so we had to try that all
clients participating in it started more or less at the same time and
from a random population. In fact, we managed to do so, except in some
byzantine cases, by making the clients restart when the cache size
collapses to 1 or to a size much smaller than the current; we called
this client configuration {\tt reboot} and {\tt cachecrash}.  This
configuration achiever a positive secondary effect: all clients
contribute to fill the cache, having thus an actual effect on the
chromosome pool and effectively making all clients present contribute
to the simulation.

The last configuration managed to find, with a relatively small number
of evaluations, solutions in half the time that in previous sets of
experiments. Although in the {\tt base} and {\tt graph} the median
time and number of contributions was smaller, we cannot disregard the
possibility that many of those solutions were found thanks to
contributions from {\em carried over} islands with an algorithm that
has run a big amount of evaluations already. At any rate, the {\tt
  cachereboot} finds on average solutions four times as fast as {\tt
  graph2} with twice as many average volunteers (or three times as
many if we consider {\em effective} volunteers), needing 50\% less
evaluations, thus meaning a big improvement in the performance of the
evolutionary algorithm, which was the target of this paper. This
improvement is less with respect to {\tt reboot}, but in any case the
method to detect the beginning of a new run is much more effective and
will be kept in the future.

>>>>

Our main intention in this paper was to configure an evolutionary
algorithm in a volunteer computing environment to improve its
efficiency and eventually find out what this efficiency depends
on. The baseline configuration that we had used in other papers  included
{\em overlaps} or volunteers running islands that had actually started in
a previous experiment instance in the set. In previous papers, we were
interested in measuring the number of users or their behavior in bulk
and this overlap was not so important, but
in this case our focus was on the evolutionary algorithm so we had to
make all volunteers in it started more or less at the same time and
from a random population. In fact, we managed to do so, except in some
byzantine cases, by making the clients restart when the cache size
collapses to 1 or to a size much smaller than the current; we called
this client configuration {\tt reboot} and {\tt cachecrash}.  This
configuration achiever a positive secondary effect: all clients
contribute to fill the cache, having thus an actual effect on the
chromosome pool and effectively making all clients present contribute
to the simulation.

The last configuration managed to find, with a relatively small number
of evaluations, solutions in half the time that in previous sets of
experiments. Although in the {\tt base} and {\tt graph} the median
time and number of contributions was smaller, we cannot disregard the
possibility that many of those solutions were found thanks to
contributions from {\em carried over} islands with an algorithm that
has run a big amount of evaluations already. At any rate, the {\tt
  cachereboot} finds on average solutions four times as fast as {\tt
  graph2} with twice as many average volunteers (or three times as
many if we consider {\em effective} volunteers), needing 50\% less
evaluations, thus meaning a big improvement in the performance of the
evolutionary algorithm, which was the target of this paper. This
improvement is less with respect to {\tt reboot}, but in any case the
method to detect the beginning of a new run is much more effective and
will be kept in the future. But the fact that this happened was due to
the {\em entropy} as proved in the last section of this paper.
